\sysfull

### 2019-07-02 paper/TOSN/coalescing.tex
Alexei feedback

\begin{algorithm}[t]
	\caption{Coalescing}
	\label{algo:genCoalescing}
	\small
%	\scriptsize
	\begin{algorithmic}[1]
        \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$} \Comment{Update coalescing target by reboot update function}
        \State $H \gets 0$ \Comment{Initialize history}
        \While{$\texttt{true}$}       
	        \State $ i \gets 0$
	        %
	        \While{$ i <= C$}
		        \State \Call{execute\_task}{$T_\texttt{i}$}
		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$} \Comment{Assign new task-dependent weight}
		        \State $i \gets i + W$
				\State $H \gets H + W$
	        \EndWhile
	        %
	        \State \Call{commit\_to\_fram}{\null}
                \State $C \leftarrow $ \Call{$f_\text{commit}$}{$C$} \Comment{Update coalescing target by commit update function}
        \EndWhile
	\end{algorithmic}
\end{algorithm}

@@

{\em Energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly. In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure. The characteristic operations of EO are
%
\begin{equation}
	\begin{split}
		f_\text{reboot}(C) & = C - x, \\
		f_\text{weight}(T_\text{i}) & =  1, \\
		f_\text{commit}(C) & = C + x.
	\end{split}
	\label{eq:eo}
\end{equation}
%
EO reacts slowly to the variation in the energy required to execute different tasks and to the variation in the effective quantum of energy available to the device. With the successful commit of each coalesced task sequence, $C$ increases. Eventually, the target may be too high and only a coalesced task composed of a few units will commit without interruption by a power failure. The strategy then linearly decreases $C$, eventually reaching a value that allows completion. A key limitation of this algorithm lies in the equality of the target decrease in $f_\text{reboot}$ and the increase in $f_\text{commit}$. Let us assume $x=1$. If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive power failures, before being able to progress.

@@

The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions. It uses its recent execution history, H, to estimate energy availability and alters its target accordingly. The EG algorithm is characterized by the following functions
%
\begin{equation}
	\begin{split}
		 f_\text{reboot}(H) & = \lceil \rho H \rceil,\\
		 f_\text{weight}(T_\text{i}) & = 1, \\
		 f_\text{commit}(C) & = \lceil \gamma C \rceil,
	\end{split}
	\label{eq:eg}
\end{equation}
%
where $\rho, \gamma \in [0, 1]$. By relying on the history of execution EG eliminates the problem of frequent power failures on a single coalesced task. In fact, at each reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
Fig.~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$. The figure shows two power cycles following the $i$th, the latter having an execution history $H^{(i)} = 4$. Based on that, $C$ is initially set to two and then decreases to one. Once the value of $C$ reaches one, EG is expecting a power failure, justifying the conservative approach. After the reboot, EG uses the most recent history $H^{(i+1)} = 8$ to set $C = 4$ and continue execution.

>>>>

\begin{algorithm}[t]
	\caption{Coalescing}
	\label{algo:genCoalescing}
	\small
%	\scriptsize
	\begin{algorithmic}[1]
        \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$} \Comment{Update coalescing target by reboot update function}
        \State $H \gets 0$ \Comment{Initialize history}
        \While{$\texttt{true}$}       
	        \State $ j \gets 0$
	        %
	        \While{$ j \leq C$}
		        \State \Call{execute\_task}{$T_\texttt{i}$} \Comment{$T_\texttt{i}$ is the $i$th task executed since the last power failure}
		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$} \Comment{Assign new task-dependent weight}
		        \State $j \gets j + W$
				\State $H \gets H + W$
	        \EndWhile
	        %
	        \State \Call{commit\_to\_fram}{\null}
                \State $C \leftarrow $ \Call{$f_\text{commit}$}{$C$} \Comment{Update coalescing target by commit update function}
        \EndWhile
	\end{algorithmic}
\end{algorithm}

@@

{\em Energy-Oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly. In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure. The characteristic operations of EO are
%
\begin{equation}
	\begin{split}
		f_\text{reboot}(C) & = C - x, \\
		f_\text{weight}(T_\text{i}) & =  1, \\
		f_\text{commit}(C) & = C + x.
	\end{split}
	\label{eq:eo}
\end{equation}
%
EO reacts slowly to the variation in the energy required to execute different tasks and to the variation in the effective quantum of energy available to the device. With the successful commit of each coalesced task sequence, $C$ increases. Eventually, the target may be too high and only a coalesced task composed of a few units will commit without interruption by a power failure. The strategy then linearly decreases $C$, eventually reaching a value that allows completion. A key limitation of this algorithm lies in the equality of the target decrease in $f_\text{reboot}$ and the increase in $f_\text{commit}$. Let us assume $x=1$. If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive reboots to progress.

@@

The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, to adhere to changes in energy conditions,  addressing a key limitation of the EO. It uses its recent execution history, H, to estimate energy availability and alters its target accordingly. The EG algorithm is characterized by the following functions
%
\begin{equation}
	\begin{split}
		 f_\text{reboot}(H) & = \lceil \rho H \rceil,\\
		 f_\text{weight}(T_\text{i}) & = 1, \\
		 f_\text{commit}(C) & = \lceil \gamma C \rceil,
	\end{split}
	\label{eq:eg}
\end{equation}
%
where $\rho, \gamma \in [0, 1]$. By relying on the history of execution EG eliminates the problem of frequent power failures on a single coalesced task. In fact, at each reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
Fig.~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$. The figure shows two power cycles following the $i$th, the latter having an execution history $H^{(i)} = 4$. Based on that, $C$ is initially set to two and then decreases to one. Once the value of $C$ reaches one, EG is expecting a power failure, justifying the conservative approach. After the reboot, EG uses the most recent history $H^{(i+1)} = 8$ to set $C = 4$ and continue execution.

### 2019-07-02 paper/TOSN/evaluation.tex
Alexei feedback

For each coalescing strategy from Section~\ref{sec:task_adaptation} (EO, EG,
and WEG) and for a baseline without coalescing (NC), we have measured the time
spent on executing (i) useful task code, (ii) task code wasted due to a power failure, and
(iii) commits to non-volatile memory at the end of each (coalesced)
task.
%
The overhead incurred by each coalescing strategy is broken down in
Fig.~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
re-execution penalty is the smallest, because the amount of work that can happen
between commits and may have to be re-executed if interrupted is reduced when
work from multiple static tasks is not combined.
%
However, any gain from a reduced re-execution penalty is canceled out by
the increased commit overhead that is incurred at the end of each static
task.
%
Across all benchmarks, all coalescing strategies reduce more commit overhead
than the re-execution overhead they add.
%
This net overhead reduction is greatest in EG and WEG strategies compared
to the EO strategy. We attribute this discrepancy to EO's slow adjustment
of the coalescing target without regard to the energy conditions.
%
In the subsequent experiments, we focus on the better-performing EG and WEG.

@@

\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{llllll}
  \toprule
    \textbf{Cap. size} ($\mu\text{f}$)& \textbf{Exp. time} ($s$)& \textbf{Duty cycle} & \textbf{Coalesced task} ($ms$)   & \textbf{Runs} & \textbf{Run time} ($ms$) \\
        \hline
    47                 &1205             & 12.93\%    & 33                        & 85       & 183      \\
    470                &1205             & 12.79\%    & 88                        & 87       & 177 \\
  \bottomrule
\end{tabular}
\end{center}
\end{minipage}  
\end{table}%

@@

Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
execution is negligible and the stored energy is insufficient to complete some of
the static tasks.}. This is marked with $\infty$ signs in Fig.~\ref{fig:runtime}. \sys, however, managed to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.

\paragraph{Different Capacitor Sizes}
Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy.
We see that \sys scales up its coalesced task size with a bigger energy buffer and vice versa. This allows it to reduce the time-to-completion of the applications. For example, 
the \textit{sort} run-time is reduced from 183\,ms to 177\,ms when the capacitor is changed from 47\,$\mu$F to 470\,$\mu$F. It should be emphasized that a device with a bigger energy buffer suffers less from power failures (but requires longer charging time). However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.

>>>>

For each coalescing strategy from Section~\ref{sec:task_adaptation} (EO, EG,
and WEG) and for a baseline without coalescing (NC), we have measured the time
spent on executing (i) useful task code, (ii) task code wasted due to a power failure, and
(iii) commits to non-volatile memory at the end of each (coalesced)
task.
%
The overhead incurred by each coalescing strategy is broken down in
Fig.~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
re-execution penalty is the smallest, because the amount of work that can happen
between commits and may have to be re-executed if interrupted is reduced when
work from multiple static tasks is not combined.
%
However, any gain from a reduced re-execution penalty is canceled out by
the increased commit overhead that is incurred at the end of each static
task.
%
Across all benchmarks, will all coalescing strategies \sys reduces more commit overhead
than the re-execution overhead they add.
%
This net overhead reduction is greatest in EG and WEG strategies compared
to the EO strategy. We attribute this discrepancy to EO's slow adjustment
of the coalescing target without regard to the energy conditions.
%
In the subsequent experiments, we focus on the better-performing EG and WEG.

@@

\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{llllll}
  \toprule
    \textbf{Cap. size} ($\mu\text{f}$)& \textbf{Exp. time} ($s$)& \textbf{on/off cycle} & \textbf{Coalesced task} ($ms$)   & \textbf{Runs} & \textbf{Run time} ($ms$) \\
        \hline
    47                 &1205             & 12.93\%    & 33                        & 85       & 183      \\
    470                &1205             & 12.79\%    & 88                        & 87       & 177 \\
  \bottomrule
\end{tabular}
\end{center}
\end{minipage}  
\end{table}%

@@

Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm
% \footnote{At distances greater than 15\,cm, energy incoming during
% execution is negligible and the stored energy is insufficient to complete some of the static tasks.}. 
\footnote{At distances less than 15\,cm the total amount of energy available for task execution includes significant amount of harvested-while-executing energy in addition to the stored energy.}.

This is marked with $\infty$ signs in Fig.~\ref{fig:runtime}. \sys, however, managed to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.

\paragraph{Different Capacitor Sizes}
Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy at runtime. This means that applications implemented in \sys are portable across devices with different capacitor sizes without recompilation; they are also more resilient to degradation in capacitor size due to temperature and device lifetime.

We see that \sys scales up its coalesced task size with a bigger energy buffer and vice versa. 
This allows it to reduce the time-to-completion of the applications. For example, 
the \textit{sort} run-time is reduced from 183\,ms to 177\,ms when the capacitor is changed from 47\,$\mu$F to 470\,$\mu$F. It should be emphasized that a device with a bigger energy buffer suffers less from power failures (but requires longer charging time). 
%However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.

### 2019-07-02 paper/TOSN/implementation.tex
minor corrections
Alexei feedback

We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittency-safe.
%
\subsection{Application Programming Interface}
\label{sec:coala_api}
%
Coala's API adds only a few syntactic constructs to a C-based language, summarized
in Table~\ref{table:coala_api}.
%
\begin{table}%
\caption{Simulation Configuration}
\label{tab:one}
\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{ll}
  \toprule
	{Method} & {Arguments} \\
	\hline
	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
	\texttt{RUN}() & --- \\
	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
	\texttt{DISABLE\_PC}() & --- \\
	\texttt{ENABLE\_PC}() & --- \\
  \bottomrule
\end{tabular}
\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
\label{table:coala_api}
\end{center}
\end{minipage}
\end{table}%
%%
%\begin{table}
%\footnotesize
%\centering
%\begin{tabular}{| r | p{0.7\columnwidth} |}
%	\hline
%	{Method} & {Arguments} \\
%	\hline\hline
%	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
%	\hline
%	\texttt{RUN}() & --- \\
%	\hline
%	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
%	\hline
%	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
%	\hline
%	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
%	\hline
%	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
%	\hline	
%	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
%	\hline
%	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
%	\hline
%	\texttt{DISABLE\_PC}() & --- \\
%	\hline
%	\texttt{ENABLE\_PC}() & --- \\
%	\hline
%\end{tabular}
%\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
%\label{table:coala_api}
%\end{table}

@@

\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed after the current one task and it can be invoked along
any control path to dynamically determine the next task.

\noindent\textit{Protected Variables.} The \texttt{PV} annotation on a
variable statically allocates a protected non-volatile variable. The
variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
run-time to ensure correct operation.

@@

\noindent\textit{Execution.} The programmer passes control to \sys's task scheduler 
by calling \texttt{RUN} after device initialization.

First, the scheduler updates the coalescing target using the coalescing
strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
the scheduler clears the list of dirty shadow pages. The scheduler then 
sets the program counter to the next task to run, which \sys tracks in  
non-volatile memory.  
%
Before executing the task, \sys checks whether there is a partially committed
task to resume, which requires \sys to restore the volatile state, including the program
counter.  
%
If there is no in-progress, partially committed task, \sys starts executing and
coalescing tasks.

@@

%
\subsection{Task Downscaling}
%
% \sys's task downscaling preserves forward progress by partially committing an
% ongoing, non-terminating task. 
%
\paragraph{Timer Strategy.}
After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state. \sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer). If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits. If the timer does not expire before power fails, \sys halves the timer and tries again. The exponential decrease of the timer value converges rapidly to a usable one.
%
%Partial commit relies on the existence of a hardware timer, which is a very common feature in a typical MCU and consumes very little energy (i.e., $\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).

@@

\paragraph{Alignment.}
Page tagging imposes a data alignment requirement.
The page size $S$ must be a power of two. 
%
Pages must be aligned to an $S$-byte boundary for efficient memory access. 
%
To preserve alignment, when typedef'ing a C structure for protected variables, the programmer has to use the API method \texttt{SM} on all members of the structure (only when defining the structure).
%
We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
an 8\,KB non-volatile \texttt{\underline{private}} buffer, and a \texttt{working} buffer
of 1\,KB.

@@

>>>>

We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittency-safe.
%
\subsection{Application Programming Interface}
\label{sec:coala_api}
%
Coala's API adds only a few syntactic constructs to a C-based language, summarized
in Table~\ref{table:coala_api}.
%
\begin{table}%
\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{ll}
  \toprule
	{Method} & {Arguments} \\
	\hline
	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
	\texttt{RUN}() & --- \\
	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
	\texttt{DISABLE\_PC}() & --- \\
	\texttt{ENABLE\_PC}() & --- \\
  \bottomrule
\end{tabular}
\label{table:coala_api}
\end{center}
\end{minipage}
\end{table}%

@@

\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed after the current one and it can be invoked along
any control path to dynamically determine the next task.

\noindent\textit{Protected Variables.} The \texttt{PV} annotation on a
variable statically allocates a protected non-volatile variable. The
variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at run-time to ensure correct operation. The \texttt{SM} annotation is used to align C structure data type within a page.

@@

\noindent\textit{Execution.} The programmer passes control to \sys's task scheduler 
by calling \texttt{RUN} after device initialization. 
\noindent\textit{Partial Commit.} The \texttt{DISABLE\_PC}() and \texttt{ENABLE\_PC}() allow a programmer to disable partial commit around certain code.

On a reboot, \sys's scheduler does a number of system level operations before executing a tast. 
 First, it updates the coalescing target according to the applied coalescing
strategy. Then, it finishes any interrupted commit and clears the list of dirty shasow pages.
% Second, the scheduler resumes a commit, if one is in progress. 
% Third, the scheduler clears the list of dirty shadow pages. 
After that, the scheduler sets the program counter to the next task to run, which \sys tracks in  
non-volatile memory.  
%
Before executing the task, \sys checks whether there is a partially committed
task to resume, which requires \sys to restore the volatile state, including the program
counter.  
%
If there is no in-progress, partially committed task, \sys starts executing and
coalescing tasks.

@@

%
\subsection{Task Downscaling}
%
% \sys's task downscaling preserves forward progress by partially committing an
% ongoing, non-terminating task. 
%
\paragraph{Timer Strategy.}
After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state. \sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer). If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits. If the timer does not expire before power fails, \sys halves the timer and tries again. The exponential decrease of the timer value converges rapidly to a usable one.

@@

\paragraph{Alignment.}
Page tagging imposes a data alignment requirement.
The page size $S$ must be a power of two. 
%
Pages must be aligned to an $S$-byte boundary for efficient memory access. 
%
To preserve alignment, when typedef'ing a C structure for protected variables, the programmer has to use the API method \texttt{SM} on all members of the structure (only when defining the structure).
%

@@

We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and an 8\,KB non-volatile \texttt{\underline{private}} buffer, and a \texttt{working} buffer
of 1\,KB.

### 2019-07-02 paper/TOSN/paging.tex
Alexei feedback

Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when tasks are coalesced. Not handling WAR dependencies on a coalesced task scope introduces a new problem, which we denote as \emph{task coalescing-induced WAR dependency}. This problem is illustrated in Fig.~\ref{figure:coalescedWar}. Therefore, \sys implements novel {\em Virtual Memory Manager} (VMM) which enables task coalescing guaranteeing data integrity, which is at the same time efficient.

VMM overview is given in Fig.~\ref{figure:mem-man}.
VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
\texttt{\underline{private}} holds the consistent version of pages after each commit (and on a reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
The VMM prohibits applications from directly accessing these buffers.
Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
When a coalesced task completes, the VMM ensures that {\em dirty}, i.e., modified, pages are committed to FRAM and visible to subsequent tasks.
If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.

@@

\subsection{Atomic Two-Phase Commit of Dirty Pages}

@@

\sys's commit is efficient because it maintains an index of dirty pages instead of iterating through all potentially dirty pages to check their state.
Another source of efficiency lies in the second phase of the commit:
% \sys's implementation is also efficient because
the VMM does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.

@@

\begin{algorithm}[t]
	\caption{Two-phase commit}
	\label{algo:commit}
	\small
	\begin{algorithmic}[1]
        \Procedure{CommitPhase1}{} \Comment{On last coalesced task}
            \For {$i \in 0..|\texttt{working}|-1$}
                \State $t \gets \Call{Tag}{\texttt{working}[i]}$
                \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
                \State $\texttt{\underline{shadowList}}[\texttt{\underline{shadowCount}}] \gets t$
                \State $\texttt{\underline{shadowCount}} \gets \texttt{\underline{shadowCount}} + 1$
            \EndFor
            \State \Call{CommitPhase2}{}
        \EndProcedure
        \Procedure{CommitPhase2}{} \Comment{\texttt{\underline{commitIndex}} 0 on first boot}
            \State $\texttt{\underline{committing}} \gets \textbf{true}$
            \While{\texttt{\underline{commitIndex}} < \texttt{\underline{shadowCount}}}
                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
                \State \Call{commitToPrivate}{\texttt{\underline{shadow}}[t]}
                \State $\texttt{\underline{commitIndex}} \gets \texttt{\underline{commitIndex}} + 1$
            \EndWhile
            \State $\texttt{\underline{shadowCount}} \gets 0$
            \State $\texttt{\underline{commitIndex}} \gets 0$
            \State $\texttt{\underline{committing}} \gets \textbf{false}$
        \EndProcedure
        \Procedure{OnBoot}{} \Comment{Invoked on every boot}
            \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
            \EndIf
            \State $\texttt{\underline{shadowCount}} \gets 0$, $i_\text{victim} \gets 0$
        \EndProcedure
	\end{algorithmic}
\end{algorithm}

>>>>

Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when tasks are coalesced. Not handling WAR dependencies on a coalesced task scope introduces a new problem, which we denote as \emph{task coalescing-induced WAR dependency}. This problem is illustrated in Fig.~\ref{figure:coalescedWar}. Therefore, \sys implements novel {\em Virtual Memory Manager} (VMM) that enables safe task coalescing and ensures efficient data manipulation.

VMM overview is given in Fig.~\ref{figure:mem-man}.
VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
\texttt{\underline{private}} holds the consistent version of pages after each commit (and on a reboot), while \texttt{\underline{shadow}} enables atomic two-phase commit: it allows \sys to ensure a persistent and consistent set of non-volatile pages before copying them to \texttt{\underline{private}} (Section~\ref{sec:two_phase_commit}).

The VMM prohibits applications from directly accessing these buffers.
Instead, it redirects any request to the \texttt{working} buffer: a relatively small buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks. With the help of \texttt{\underline{shadow}}, the \texttt{working} buffer can serve more pages than its own capacity.

When a coalesced task completes, the VMM ensures that {\em dirty}, i.e., modified, pages are committed to FRAM and visible to subsequent tasks.
If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.

@@

\subsection{Atomic Two-Phase Commit of Dirty Pages}
\label{sec:two_phase_commit}

@@

\sys's commit is efficient because it maintains an index of dirty pages instead of iterating through all potentially dirty pages to check their state.
Another source of efficiency lies in the second phase of the commit:
% \sys's implementation is also efficient because
the VMM does not copy page content from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.

@@

\begin{algorithm}[t]
	\caption{Two-phase commit}
	\label{algo:commit}
	\small
	\begin{algorithmic}[1]
        \Procedure{CommitPhase1}{} \Comment{On completion of a coalesced task}
            \For {$i \in 0..|\texttt{working}|-1$}
                \State $f_\text{dirty}, t \gets \Call{TagFlag}{\texttt{working}[i]}$
                % \State $f_\text{dirty} \gets \Call{DirtyFlag}{t}$
                \If {$f_\text{dirty}$}   \Comment{Is the page content modified?}
                    \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
                    \State $\texttt{\underline{shadowList}}[\texttt{\underline{shadowCount}}] \gets t$
                    \State $\texttt{\underline{shadowCount}} \gets \texttt{\underline{shadowCount}} + 1$
                \EndIf
            \EndFor
            \State \Call{CommitPhase2}{}
        \EndProcedure
        \Procedure{CommitPhase2}{} \Comment{\texttt{\underline{commitIndex}} 0 on first boot}
            \State $\texttt{\underline{committing}} \gets \textbf{true}$
            \While{\texttt{\underline{commitIndex}} < \texttt{\underline{shadowCount}}}
                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
                \State \Call{commitToPrivate}{\texttt{\underline{shadow}}[t]} \Comment{Copy \texttt{\underline{shadow}} to \texttt{\underline{private}} by swapping their pointers}
                \State $\texttt{\underline{commitIndex}} \gets \texttt{\underline{commitIndex}} + 1$
            \EndWhile
            \State $\texttt{\underline{shadowCount}} \gets 0$
            \State $\texttt{\underline{commitIndex}} \gets 0$
            \State $\texttt{\underline{committing}} \gets \textbf{false}$
        \EndProcedure
        \Procedure{OnBoot}{} \Comment{Invoked on every boot}
            \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
            \EndIf
            \State $\texttt{\underline{shadowCount}} \gets 0$ %, $i_\text{victim} \gets 0$
        \EndProcedure
	\end{algorithmic}
\end{algorithm}

### 2019-07-02 paper/TOSN/partialcommit.tex
First iteration
Alexei feedback

\sys uses {\em task downscaling} to make progress through a task that is too large to complete using the energy that the device can buffer. Task downscaling executes part of the long task and interrupts its execution with a {\em partial commit}, after which the long task continues. This timer-based solution is similar to the one proposed in~\cite{ratchet}. If power fails, the partial commit preserves the progress through the task, and execution resumes from the point of the partial commit, rather than from the start of the task. Eventually, after some number of partial commits and power failures, the task will complete and execution will proceed. Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity. If a programmer requires a task's atomicity to be preserved, they can disable task downscaling for that task. The key design issue for task downscaling is deciding when to partially commit. A task is likely to be non-terminating if it fails to run to completion twice consecutively. The second incomplete run executes after a power failure, when the device will have fully recharged its energy buffer. If a task cannot complete when executing with a full energy buffer, \sys marks the task as non-terminating.

>>>>

\sys uses {\em task downscaling} to make progress through a task that is too large to complete using the buffered energy. Task downscaling executes part of the long task and interrupts its execution with a {\em partial commit}, after which the long task continues. This timer-based solution is similar to the one proposed in~\cite{ratchet}. If power fails, the partial commit preserves the progress through the task, and execution resumes from the point of the partial commit, rather than from the start of the task. Eventually, after some number of partial commits and power failures, the task will complete and execution will proceed.

\textit{Detection of Non-terminating Tasks.} The key design issue for task downscaling is deciding when to partially commit. A task is likely to be non-terminating if it fails to run to completion twice consecutively. 
The second incomplete run executes after a power failure, when the device will have fully recharged its energy buffer. If a task cannot complete when executing with a full energy buffer, \sys marks the task as non-terminating.
Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity. If a programmer requires a task's atomicity to be preserved, they can disable task downscaling for that task.

### 2019-07-02 paper/TOSN/relatedwork.tex
Alexei feedback

In all of the above systems, with the exception of DINO~\cite{dino}, checkpoints are dynamic, i.e., the programmer does not have explicit control over the point at which the code may be resumed after power failure. Dynamic checkpointing systems make it difficult for the programmer to respect application-level atomicity constraints, such as correlating sensor readings. Checkpointing systems that copy most volatile state scale poorly as the size of volatile memory increases.  QuickRecall, Clank, and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf. Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile memory. In addition, QuickRecall and Clank require custom hardware.

>>>>

In all of the above systems, with the exception of DINO~\cite{dino}, checkpoints are dynamic, i.e., the programmer does not have explicit control over the point at which the code may be resumed after power failure. Dynamic checkpointing systems make it difficult for the programmer to respect application-level atomicity constraints, such as correlating sensor readings. Checkpointing systems that copy most volatile state scale poorly as the size of volatile memory increases.  QuickRecall, Clank, and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory, whichrequires more time and energy to access than volatile memory (cf. Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile memory. In addition, QuickRecall and Clank require custom hardware.
